Predicting complex diseases: performance and robustness

### 2017-10-26 paper/draft.tex
save

@@

\subsection{Simulations}

@@

@@

@@

@@

>>>>

\let\tabbeg\tabular
\let\tabend\endtabular
\renewenvironment{tabular}{\begin{adjustbox}{max width=\textwidth}\tabbeg}{\tabend\end{adjustbox}}

@@

\subsection{Simulations} \label{sec:simus}

@@

In this study, we use two different measures of predictive accuracy. 
First, we use the Area Under the Receiver Operating Characteristic (ROC) Curve (AUC). In the case of our study, the AUC is the probability that a PRS of a case is greater than one of a control.
This measure indicates how well we can distinguish between cases and controls using PRSs.
As a second measure, we report the percentage of cases in the 10\% largest PRSs [UTILISER 20\%?]. This measure indicates how well we can identify people at higher risk from a higher PRS in screening campaigns.

Note that we also report the timing of the main computations and the number of SNPs used for the prediction.

@@

\subsection{Reproduciblity}

All the code used in this paper along with results,  such as execution times and figures,  are available as HTML R notebooks in the Supplementary Materials.

@@

\subsection{Overview and naming}

\subsection{T-Trees}

First, we ran only 5 simulations for each combination of parameters (see section \ref{sec:simus}), excepted for the heritability that we fixed at $0.8$.
The goal was to quickly discard the T-Trees method as being competitive with the other methods. 
For example, compared to our penalized logistic regression, T-Trees perform worse for both predictive measures, the AUC and the percentage of cases in highest scores. Moreover, T-Trees takes longer to run and makes more complex predictive models because it uses more SNPs in the models and have non-linear effects (Figure \ref{fig:ttrees}).
So, we decided to discard this method for the rest of the simulations.

\subsection{Main simulations}

For the P+T method, we report three different scores of prediction: one including all the SNPs (after clumping), one including only SNPs that have a p-value under the GWAS threshold of significance ($p < 5 \cdot 10^{-8}$), and one that maximize the AUC for these two thresholds and a sequence of 100 values from $10^{-0.1}$ to $10^{-100}$ equally spaced on the log-log-scale (Table \ref{tab:thr}). Note that, normally, the optimal threshold would have to be learned on the training set only. We consider this reported maximum AUC as an upper bound of the AUC for the P+T method.

[Also return p-values?]

[TIMING: PRS -> 0; logit -> 500-600 with small increase when more causal variants, logit-triple x3]

[AUC: decrease with M, PRS-all never good, logit-triple better for fancy and almost as good for simple]

[\% Cases: Linear with AUC, Attention perc cases is == 30%]

\section{Discussion}

@@

\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth]{ttrees}}
\caption{Results of T-Trees vs penalized logistic regression. \textbf{A.} Timing (in seconds). \textbf{B.} Number of predictors of the model. \textbf{C.} AUC. \textbf{D.} Percentage of cases in the 10\% largest scores.}\label{fig:ttrees}
\end{figure}

% latex table generated in R 3.4.1 by xtable 1.8-2 package
% Thu Oct 26 16:46:26 2017
\begin{table}[ht]
\centering
\begin{tabular}{ccccccccccccccc}
  \hline
1.00e+00 & 7.22e-01 & 5.87e-01 & 4.20e-01 & 2.43e-01 & 1.00e-01 & 2.35e-02 & 2.21e-03 & 4.69e-05 & 8.81e-08 & 3.18e-12 & 1.83e-19 & 2.89e-31 & 1.70e-50 & 7.71e-82 \\ 
  5.00e-08 & 7.05e-01 & 5.65e-01 & 3.95e-01 & 2.20e-01 & 8.47e-02 & 1.79e-02 & 1.42e-03 & 2.28e-05 & 2.73e-08 & 4.69e-13 & 8.08e-21 & 1.80e-33 & 4.30e-54 & 1.06e-87 \\ 
  7.94e-01 & 6.87e-01 & 5.42e-01 & 3.69e-01 & 1.97e-01 & 7.08e-02 & 1.34e-02 & 8.83e-04 & 1.05e-05 & 7.74e-09 & 6.03e-14 & 2.86e-22 & 7.73e-36 & 5.97e-58 & 5.49e-94 \\ 
  7.81e-01 & 6.69e-01 & 5.19e-01 & 3.43e-01 & 1.75e-01 & 5.85e-02 & 9.79e-03 & 5.31e-04 & 4.61e-06 & 2.01e-09 & 6.69e-15 & 7.92e-24 & 2.24e-38 & 4.37e-62 & 1.00e-100 \\ 
  7.67e-01 & 6.50e-01 & 4.95e-01 & 3.18e-01 & 1.54e-01 & 4.76e-02 & 7.01e-03 & 3.08e-04 & 1.90e-06 & 4.72e-10 & 6.32e-16 & 1.70e-25 & 4.26e-41 & 1.61e-66 &  \\ 
  7.53e-01 & 6.30e-01 & 4.70e-01 & 2.93e-01 & 1.35e-01 & 3.82e-02 & 4.90e-03 & 1.72e-04 & 7.31e-07 & 1.00e-10 & 5.04e-17 & 2.75e-27 & 5.16e-44 & 2.83e-71 &  \\ 
  7.38e-01 & 6.09e-01 & 4.46e-01 & 2.68e-01 & 1.17e-01 & 3.02e-02 & 3.33e-03 & 9.18e-05 & 2.63e-07 & 1.89e-11 & 3.35e-18 & 3.31e-29 & 3.84e-47 & 2.26e-76 &  \\ 
   \hline
\end{tabular}
\caption{The 102 thresholds used in the P+T method for this study.}
\label{tab:thr}
\end{table}
